---
title: "Heuristic Analysis"
author: "Artem Larionov"
date: "3/19/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Statistics

```{r, echo=FALSE}
df <- data.frame(rbind(
  c('Problem 1','breadth_first_search',43,56,180,6,0.022),
  c('Problem 1','depth_first_graph_search',12,13,48,12,0.006),
  c('Problem 1','uniform_cost_search',55,57,224,6,0.027),
  c('Problem 1', 'A* h_1',55,57,224,6,0.028),
  c('Problem 1', 'A* h_ignore_preconditions',41,43,170,6,0.022),
  c('Problem 1', 'A* h_pg_levelsum',55,57,224,6,0.75),
  
  c('Problem 2','breadth_first_search',3346,4612,30534,9,10),
  c('Problem 2','depth_first_graph_search',1124,1125,10017,1085,6.86),
  c('Problem 2','uniform_cost_search',4852,4854,44030,9,5.98),
  c('Problem 2', 'A* h_1',4852,4854,44030,9,5.97),
  c('Problem 2', 'A* h_ignore_preconditions',1450,1452,13303,9,1.9),
  c('Problem 2', 'A* h_pg_levelsum',4852,4854,44030,9,66),
  
  c('Problem 3','breadth_first_search',14120,17673,124926,12,81),
  c('Problem 3','depth_first_graph_search',677,678,5608,660,2.6),
  c('Problem 3','uniform_cost_search',18235,18237,159716,12,24.86),
  c('Problem 3', 'A* h_1',18235,18237,159716,12,24.6),
  c('Problem 3', 'A* h_ignore_preconditions',5040,5042,44944,12,7.2),
  c('Problem 3', 'A* h_pg_levelsum',18235,18237,159716,12,303)))

colnames(df) <- c('problem', 'search', 'Expansions', 'Goal_Tests', 'New_Nodes', 'Plan_Length', 'Time')
df$Expansions <- as.numeric(as.character(df$Expansions))
df$Goal_Tests <- as.numeric(as.character(df$Goal_Tests))
df$New_Nodes <- as.numeric(as.character(df$New_Nodes))
df$Plan_Length <- as.numeric(as.character(df$Plan_Length))
df$Time <- as.numeric(as.character(df$Time))

knitr::kable(df[order(df$problem),])
```

#### Conclusions

* Depth-first Graph Search can be fast but usually it finds not optimal plan. 
* Breadth-first Search finds an optimal plan, but explore a lot of nodes, which leads to slower execution time.
* The faster speed of DFS comparing to BFS is due to less number of expansions and nodes  tested, it's also the reason why DFS can't guarantee the optimality of the solution, which BFS can.
* Uniform Cost Search, A\* with h_1 and h_pg_levelsum heuristics have almost the same metrics. 
* A\* with h_pg_levelsum is the slowest algorithm among all variants though.
* A\* with h_ignore_preconditions heuristic has the best metrics and finds an optimal plan.
* h_ignore_preconditions allows A\* to minimize number of expansions and tests, hence to find an optimal solution much faster than other heuristics and uninform search algorithms.

The drawback of depth-first search is that it can get stuck going down the wrong path. Depth-first search will either get stuck in an infinite loop and never return a solution, or it may eventually find a solution path that is longer than the optimal solution.[^1]

If there is a solution, breadth-first search is guaranteed to find it, and if there are several solutions, breadth-first search will always find the shallowest goal state first. In terms of the four criteria, breadth-first search is complete, and it is optimal provided the path cost is a nondecreasing function of the depth of the node.
The memory requirements are a bigger problem for breadth-first search than the execution time.[^2]

That A\* search is complete, optimal, and optimally efficient among all such algorithms is rather satisfying. Unfortunately, it does not mean that A\* is the answer to all our searching needs. It has been shown that exponential growth will occur unless the error in the heuristic function grows no faster than the logarithm of the actual path cost.[^3]

[^1]: AIMA, Chapter 3. Solving Problems by Searching, Depth-first search
[^2]: AIMA, Chapter 3. Solving Problems by Searching, Breadth-first search
[^3]: AIMA, Chapter 4. Informed Search Methods, Proof of the optimality of A\*

#### Path Plans

*Problem 1*

| Breadth First Search  | Depth First Graph Search   | Uniform Cost Search  
------------------------|----------------------------|---------------------
| Load(C2, JFK, P2)     | Fly(P1, SFO, JFK)          | Load(C1, SFO, P1)   
| Load(C1, SFO, P1)     | Fly(P2, JFK, SFO)          | Load(C2, JFK, P2)   
| Fly(P2, JFK, SFO)     | Load(C1, SFO, P2)          | Fly(P1, SFO, JFK)   
| Unload(C2, P2, SFO)   | Fly(P2, SFO, JFK)          | Fly(P2, JFK, SFO)   
| Fly(P1, SFO, JFK)     | Fly(P1, JFK, SFO)          | Unload(C1, P1, JFK) 
| Unload(C1, P1, JFK)   | Unload(C1, P2, JFK)        | Unload(C2, P2, SFO) 
|                       | Fly(P2, JFK, SFO)          |                    
|                       | Fly(P1, SFO, JFK)          |                     
|                       | Load(C2, JFK, P1)          |                     
|                       | Fly(P2, SFO, JFK)          |                     
|                       | Fly(P1, JFK, SFO)          |                     
|                       | Unload(C2, P1, SFO)        |                     
                      
| A\* h_1             | A\* h_ignore_preconditions | A\* h_pg_levelsum   
----------------------|----------------------------|---------------------
| Load(C1, SFO, P1)   | Load(C1, SFO, P1)          | Load(C1, SFO, P1)   
| Load(C2, JFK, P2)   | Fly(P1, SFO, JFK)          | Load(C2, JFK, P2)   
| Fly(P1, SFO, JFK)   | Unload(C1, P1, JFK)        | Fly(P1, SFO, JFK)   
| Fly(P2, JFK, SFO)   | Load(C2, JFK, P2)          | Fly(P2, JFK, SFO)   
| Unload(C1, P1, JFK) | Fly(P2, JFK, SFO)          | Unload(C1, P1, JFK) 
| Unload(C2, P2, SFO) | Unload(C2, P2, SFO)        | Unload(C2, P2, SFO) 

*Problem 2*

 Breadth First Search | Depth First Graph Search | Uniform Cost Search 
----------------------|:------------------------:|---------------------
| Load(C1, SFO, P1)     | 1085 steps               | Load(C1, SFO, P1)
| Load(C2, JFK, P2)     |                          | Load(C2, JFK, P2)
| Load(C3, ATL, P3)     |                          | Load(C3, ATL, P3)
| Fly(P1, SFO, JFK)     |                          | Fly(P1, SFO, JFK)
| Unload(C1, P1, JFK)   |                          | Fly(P2, JFK, SFO)
| Fly(P2, JFK, SFO)     |                          | Fly(P3, ATL, SFO)
| Unload(C2, P2, SFO)   |                          | Unload(C3, P3, SFO)
| Fly(P3, ATL, SFO)     |                          | Unload(C2, P2, SFO)
| Unload(C3, P3, SFO)   |                          | Unload(C1, P1, JFK)

 A\* h_1            | A\* h_ignore_preconditions | A\* h_pg_levelsum
--------------------|----------------------------|---------------------
| Load(C1, SFO, P1)   | Load(C3, ATL, P3)          | Load(C3, ATL, P3)
| Load(C2, JFK, P2)   | Fly(P3, ATL, SFO)          | Fly(P3, ATL, SFO)
| Load(C3, ATL, P3)   | Unload(C3, P3, SFO)        | Unload(C3, P3, SFO)
| Fly(P1, SFO, JFK)   | Load(C2, JFK, P2)          | Load(C1, SFO, P1)
| Fly(P2, JFK, SFO)   | Fly(P2, JFK, SFO)          | Load(C2, JFK, P2)
| Fly(P3, ATL, SFO)   | Unload(C2, P2, SFO)        | Fly(P2, JFK, SFO)
| Unload(C3, P3, SFO) | Load(C1, SFO, P1)          | Unload(C2, P2, SFO)
| Unload(C2, P2, SFO) | Fly(P1, SFO, JFK)          | Fly(P1, SFO, JFK)
| Unload(C1, P1, JFK) | Unload(C1, P1, JFK)        | Unload(C1, P1, JFK)

*Problem 3*

Breadth First Search | Depth First Graph Search | Uniform Cost Search 
----------------------|:-----------------------:|---------------------
| Load(C1, SFO, P1)     |  660 steps              | Load(C1, SFO, P1)  
| Load(C2, JFK, P2)     |                         | Load(C2, JFK, P2)
| Fly(P1, SFO, ATL)     |                         | Fly(P1, SFO, ATL)
| Load(C3, ATL, P1)     |                         | Load(C3, ATL, P1)
| Fly(P2, JFK, ORD)     |                         | Fly(P2, JFK, ORD)
| Load(C4, ORD, P2)     |                         | Load(C4, ORD, P2)
| Fly(P1, ATL, JFK)     |                         | Fly(P2, ORD, SFO)
| Unload(C1, P1, JFK)   |                         | Fly(P1, ATL, JFK)
| Unload(C3, P1, JFK)   |                         | Unload(C4, P2, SFO)
| Fly(P2, ORD, SFO)     |                         | Unload(C3, P1, JFK)
| Unload(C2, P2, SFO)   |                         | Unload(C2, P2, SFO)
| Unload(C4, P2, SFO)   |                         | Unload(C1, P1, JFK)

 A\* h_1            | A\* h_ignore_preconditions | A\* h_pg_levelsum
--------------------|----------------------------|---------------------
| Load(C1, SFO, P1)   | Load(C2, JFK, P2)          | Load(C1, SFO, P1)
| Load(C2, JFK, P2)   | Fly(P2, JFK, ORD)          | Load(C2, JFK, P2)
| Fly(P1, SFO, ATL)   | Load(C4, ORD, P2)          | Fly(P1, SFO, ATL)
| Load(C3, ATL, P1)   | Fly(P2, ORD, SFO)          | Load(C3, ATL, P1)
| Fly(P2, JFK, ORD)   | Unload(C4, P2, SFO)        | Fly(P2, JFK, ORD)
| Load(C4, ORD, P2)   | Load(C1, SFO, P1)          | Load(C4, ORD, P2)
| Fly(P2, ORD, SFO)   | Fly(P1, SFO, ATL)          | Fly(P2, ORD, SFO)
| Fly(P1, ATL, JFK)   | Load(C3, ATL, P1)          | Unload(C4, P2, SFO)
| Unload(C4, P2, SFO) | Fly(P1, ATL, JFK)          | Fly(P1, ATL, JFK)
| Unload(C3, P1, JFK) | Unload(C3, P1, JFK)        | Unload(C3, P1, JFK)
| Unload(C2, P2, SFO) | Unload(C2, P2, SFO)        | Unload(C2, P2, SFO)
| Unload(C1, P1, JFK) | Unload(C1, P1, JFK)        | Unload(C1, P1, JFK)